---
title: "CogScie eye-tracking lab Fall 2022"
subtitle: "Data cleaning and sanity checks"
author: "Fabio Trecca"
date: "8/12/2022"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Load in the Visual Search experiment data
```{r}
Samples <- 
Fixations <- 
Saccades <- 
```

# Sanity checks

The first thing we want to do is to run some sanity checks on the sample data (x and y coordinates) to make sure everything is in order. Notice anything wrong?

```{r}
## Check distribution of raw samples. Is everything alright?

```

If needed, remove artifacts and other "illegal" data points. The screen resolution at COBE Lab was 1050 x 1680. Any points outside this range should be excluded. Start by figuring out how frequent these points are. Eg. are some participants more susceptible to outliers than others?

```{r}

```

Then filter out these outlier points:

```{r}

```


## Inspect distribution of no. of fixations and fixation duration by participant/trial both VISUALLY and NUMERICALLY

```{r}

```

What do you notice? Make sure that nothing looks funky in the data!

## Inspect distribution of no. of saccades and saccade amplitude by participant/trial both VISUALLY and NUMERICALLY

```{r}

```

What do you notice? Make sure that nothing looks funky in the data!

## Inspect distribution of pupil size by participant/trial both VISUALLY and NUMERICALLY

```{r}

```

What do you notice? Make sure that nothing looks funky in the data!

## Check distribution of Time variable. Is there anything weird?

```{r}

```

## Let's also do some outlier checks:

```{r}
## Check for outliers in fixation duration using a boxplot

```

```{r}
## Check for outliers in saccade length using a boxplot

```

