---
title: "CogSci eye-tracking lab Fall 2022"
subtitle: "Data visualization exercise"
author: "Fabio Trecca"
date: "8/12/2022"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
pacman::p_load(tidyverse, jpeg, grid, ggpubr, ggrepel)
```

## Load the data (if it's not already in your Environment)

```{r load data}

```

## Import picture information (ForagingOrder.csv) and merge with data sets

```{r}
Pictures <- 
Samples <- 
Fixations <- 
Saccades <- 
```

## Flip the data on the y-axis since it's mirrored

```{r}
Samples <- Samples %>% mutate(PositionY = 1680-PositionY)
Fixations <- Fixations %>% mutate(PositionY = 1680-PositionY)
Saccades <- Saccades %>% mutate(PositionY = 1680-PositionY)
```

Make sure that everything worked as intended.

### Scanpaths

A classic way of plotting eye tracking data is with *scanpaths*. Scanpaths show fixations as circles of different size (where larger circles mean longer fixations) and saccades as lines connecting the circles.

Let's make a scanpath for one of the pictures in the Visual foraging experiment. We'll make this into a function so that we don't have to rewrite the many lines of code every time:

```{r}
make_scanpath <- function(participant, trial){
  
  ## Select only the participant/trial of interest
  x <- Fixations[Fixations$ParticipantID == participant & Fixations$Trial == trial,]
  
  ## Use these lines to load the picture and turn in into a rasterized grob that we can plot in ggplot
  img <- readJPEG(paste0('stimuli_Foraging/', x$Stimulus[1])) 
  img <- rasterGrob(img, width=unit(1, "npc"), height = unit(1,"npc"), interpolate = FALSE)
  
  p <-
    # basically, we make a scatterplot with x and y coordinates on the x and y axis respectively
    ggplot(x, aes(PositionX, PositionY, color = FixationNo)) + 
    # plot the picture in the background
    annotation_custom(img, xmin = 0, xmax = 1680, ymin = 0, ymax = 1050) +
    # hacky way to adjust opacity of background picture:
    annotate(geom = "rect", xmin = 0, xmax = 1680, ymin = 0, ymax = 1050, fill = "white", alpha = .3) +
    # create saccade path between points
    geom_path(color = "black") +
    # create fixation points, with size as a function of duration
    geom_point(aes(size = Duration), alpha = .8) +
    # repel the labels to increase readability
    geom_text_repel(aes(label = Duration), size = 3, color = "white") +
    # make everything look nicer
    xlim(0,1680) + ylim(0,1050) +
    theme_light()
  
  print(p)
}
```

One disadvantage of scanpaths is that they are only informative for one participant in one trial at the time. As you can see, the function above requires that you select one participant and one trial.

Let's make a scanpath for one trial in the Search condition:

```{r}

```

And here is one from the Count condition:

```{r}

```

Smukt!

*Exercise:*

Make a scanpath for all 10 trials from one participant of your choice (perhaps in a loop?). Use these plots as diagnostic tools in order to answer the following questions:

- Do the data look reasonable and of good quality, or are there artefacts?

- Can we differentiate between the two conditions (Count and Search) only by looking at the scanpaths?

- Can we spot the trials in which the participant found the star?

```{r}

```

### Heatmaps

Now we are going to apply the same plotting principles as in the scanpaths, but instead of plotting actual fixations (and their interconnecting lines) we will plot a density distribution of fixations across a trial, thereby losing the temporal aspect of scanpaths (but gaining in readability since now we can visualize data from multiple participants/trials at once).

Let's choose some heatmapy colors first:

```{r}
heat_colors <- colorRampPalette(
  c(
    "#00007F",
    "blue",
    "#007FFF",
    "cyan",
    "#7FFF7F",
    "yellow",
    "#FF7F00",
    "red",
    "#7F0000"
  )
)
```

Then, we build a function that plots the heatmap using ggplot2. Notice how most lines are analogous to make_scanpath() above.

```{r}
make_heatmap <- function(participant, trial){
  
  x <- Fixations[Fixations$ParticipantID == participant & Fixations$Trial == trial,]
  
  img <- jpeg::readJPEG(paste0('pictures/', x$Stimulus[1])) 
  img <- grid::rasterGrob(img, width=unit(1, "npc"), height = unit(1,"npc"), interpolate = FALSE)
  
  p <- ggplot(Fixations, aes(PositionX, PositionY, color = FixationNo)) + 
    annotation_custom(img, xmin = 0, xmax = 1680, ymin = 0, ymax = 1050) +
    annotate(geom = "rect", xmin = 0, xmax = 1680, ymin = 0, ymax = 1050, fill = "white", alpha = .3) +
    # instead of plotting circles, we plot a density function, so that areas with different data density levels
    # have different colors. The sqrt(sqrt()) thing is just to get the ratio of warm/cold colors right
    stat_density2d(geom = "raster", aes(fill = ..density.., alpha = sqrt(sqrt(..density..))), contour = FALSE, n = 1000) +
    # we take the density data and apply the colors that we chose
    scale_fill_gradientn(colours = heat_colors(10), trans="sqrt") +
    # and apply dynamic alpha to it so that the fewer the fixations, the more transparent the color
    scale_alpha(range = c(0.1, 0.6)) + xlim(0,1680) + ylim(0,1050) +
    theme(legend.position = "none") +
    theme_light()
  
  print(p)
}
```

And now we can finally make our heatmap -- pick a participant and a trial:

```{r}

```

And one more:

```{r}

```

Wonderful!

*Exercise*:

Borrowing from the heatmap code above and readapting it, make a cumulative heatmap for all participants in the Foraging experiment looking at the ‘penguins.jpg’ image. Facet the graph by condition (Search vs. Count). What do you notice?

```{r}

```

### Areas of Interest (AOIs)

In order to draw AOIs around our stimuli, we have to figure out what their coordinates should be. This may require some trial-and-error. In this example, I want to draw a rectangle around the black sheep in the foreground. The approximate coordinates are the following:

```{r}
## Define an AOI for the black sheep
AOI = c(720, 930, 50, 330)
      #xmin xmax ymin ymax
```

Before plotting the data, I filter all the data that do not fit within the coordinates described above.

```{r}
x <-
  Fixations %>% filter(PositionX >= AOI[1] &
                         PositionX <= AOI[2] &
                         PositionY >= AOI[3] & PositionY <= AOI[4])
```

```{r}
img <- jpeg::readJPEG('pictures/sheep.jpg')  
img <- grid::rasterGrob(img, width=unit(1, "npc"), height = unit(1,"npc"), interpolate = FALSE)

ggplot(x, aes(PositionX, PositionY, color = FixationNo)) + 
  annotation_custom(img, xmin = 0, xmax = 1680, ymin = 0, ymax = 1050) +
  # this line draws the rectangle for the AOI
  geom_rect(xmin=AOI[1], xmax=AOI[2], ymin=AOI[3], ymax=AOI[4], fill = NA, size = 1, color = 'red') +
  stat_density2d(geom = "raster", aes(fill = ..density.., alpha = sqrt(sqrt(..density..))), contour = FALSE, n = 1000) + 
  scale_fill_gradientn(colours = heat_colors(10), trans="sqrt") +
  scale_alpha(range = c(0.1, 0.6)) +
  xlim(0,1680) + ylim(0,1050) +
  theme(legend.position = "none")
```

*Excercise*:

Make a cumulative heatmap for all participants looking at the ‘dolphins.jpg’ image and facet the graph by task (Search vs. Count) after having created an AOI at a location of your choice. What do you notice?

```{r}

```

## Growth curves

Growth curves show how proportional looking at one or more specific AOIs changes (not necessarily grows, despite the name…) over time and across participants. Let’s start by defining two AOIs:

```{r}
AOI1 = c(300, 700, 200, 450)
AOI2 = c(600, 1100, 600, 750)
```

Let’s make a summary dataset for fixations and filter the fixations that fall within one of the two AOIs. The plot below shows what the two AOIs look like:

```{r}
x <- Fixations %>%
  filter(Stimulus == "trees.jpg") %>%
  mutate(
    InAOI1 = ifelse(
      PositionX >= AOI1[1] &
        PositionX <= AOI1[2] &
        PositionY >= AOI1[3] & PositionY <= AOI1[4],
      TRUE,
      FALSE
    ),
    InAOI2 = ifelse(
      PositionX >= AOI2[1] &
        PositionX <= AOI2[2] &
        PositionY >= AOI2[3] & PositionY <= AOI2[4],
      TRUE,
      FALSE
    )
  )

img <- jpeg::readJPEG('pictures/trees.jpg')  
img <- grid::rasterGrob(img, width=unit(1, "npc"), height = unit(1,"npc"), interpolate = FALSE)

ggplot(x, aes(PositionX, PositionY, color = FixationNo)) + 
  annotation_custom(img, xmin = 0, xmax = 1680, ymin = 0, ymax = 1050) +
  geom_rect(xmin=AOI1[1], xmax=AOI1[2], ymin=AOI1[3], ymax=AOI1[4], fill = NA, size = 1, color = 'red') +
  annotate(geom = "label", x = 500, y = 450, label = "Birds", color = "red") +
  geom_rect(xmin=AOI2[1], xmax=AOI2[2], ymin=AOI2[3], ymax=AOI2[4], fill = NA, size = 1, color = 'blue') +
  annotate(geom = "label", x = 850, y = 750, label = "More birds", color = "blue") +
  xlim(0,1680) + ylim(0,1050) +
  theme(legend.position = "none") +
  theme_light()
```

Now let’s make a new summary dataset where we compute proportions of fixations in either of the two AOIs divided by total number of fixations, and let’s plot this proportion using a smoothing function. Which area was fixated first?

```{r}
Prop <- x %>% 
  group_by(FixationNo) %>% 
  summarize(AOI1 = sum(InAOI1 == TRUE)/(length(InAOI1)+length(InAOI2))*100,
            AOI2 = sum(InAOI2 == TRUE)/(length(InAOI1)+length(InAOI2))*100) %>% 
  gather("AOI", "Proportion", AOI1:AOI2)

ggplot(Prop, aes(FixationNo, Proportion, color = AOI)) +
  geom_smooth() + ylim(-10,100) +
  theme_light()
```

*Exercise*:

Try adding a third AOI and computing proportional looks to it like we did above.

```{r}


```

*Exercise*:

In the sheep picture, draw an AOI around the black sheep and compute the average time to first fixation (TTFF) to that AOI. What do you notice?

```{r}


```
